Macroscopic Salt Rejection through Electrostatically Gated Porous Graphene

### 2018-12-22 paper/paper.tex
Implement manuscript changes with Roman

\author{Roman M. Wyss} \affiliation{Institute of Soft Materials
  Department of Material Sciences, Eidgenössische Technische
  Hochschule (ETH) Zürich, Vladimir-Prelog-Weg 1-5, Zürich CH-8093,
  Switzerland.  }
\altaffiliation{R. M. S. and T. T. contributed equally to this work}
\author{Tian Tian} \affiliation{Institute for
  Chemical and Bioengineering Department of Chemistry and Applied
  Biosciences, Eidgenössische Technische Hochschule (ETH) Zürich,
  Vladimir-Prelog-Weg 1-5, Zürich CH-8093, Switzerland.}
\altaffiliation{R. M. S. and T. T. contributed equally to this work}

@@

%%%%%%%%%%%%%%%%%%%%
%Main
%%%%%%%%%%%%%%%%%%%%
\begin{document}
\clearpage{}
\begin{abstract}
  {\normalsize Atomically thin graphene nanopores is emerging as one
    of the most promising candidates for next-generation membrane
    technology owing to the ultrahigh permeability. However, the
    size-exclusive mechanism of selectivity limits its scalability as
    it remains technically challenging to drill subnanometer pores
    over a large area. Here we report macroscopic salt rejection
    through atomically-thin porous graphene with the pore sizes of
    20$\pm$10 nanometers, and a pore density of 1.25$\pm$0.25$\times$10$^{10}$
    cm$^{-2}$, overcoming the pore size limitation set by the hydrated
    radii. By electrostatically gating a sheet of porous graphene, we
    report a considerable reduction of salt flux by up to 55\% and
    85\% with the gate voltage, based on our experiments and
    theoretical predictions, respectively. We systematically
    investigate the effects of salt concentration and species,
    including developing a theory to model the electrolyte diffusion
    through a nanopore in a sheet of gated graphene. The interplay
    between graphene quantum capacitance and electrical double layer
    is identified to be the main mechanism responsible for the
    observed salt rejection, when the pore size is comparable to the
    Debye length. Our findings reveal a new degree of
    freedom controlling the membrane potential and electrolyte
    permeation through porous two-dimensional materials.  }
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Main text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage{}
\linenumbers
\section{Introduction}
\label{sec:intro}
A central objective in the emerging field of nanofluidics is to
manipulate transport phenomena at nanometer scale to enable new
technological opportunities. The reduction of characteristic
dimensions in nanofluidics gives rise to substantially different
transport properties in comparison to their bulk counterparts
\cite{Schoch_2008}. In nanopores, for example, the effect of
surface-mediated transport becomes dominant when the pore size is
comparable to the characteristic length of diffusion, enabling new
applications such as ionic diodes \cite{Karnik_2007}, field-effect
transistors \cite{Nam_2009}, desalination \cite{Heiranian_2015}, and
nanopore-based DNA sequencing \cite{Heerema_2016,Garaj_2013}, Porous,
atomically thin two-dimensional (2D) materials, such as graphene, are
emerging as ideal membrane materials due to their ultimate permeation
\cite{Suk_2010,Jiang_2009,Celebi_2014,Koenig_2012,Drahushuk_2012}. Early
research has suggested that using nanoporous graphene as the
desalination membrane enables excellent salt rejection by preventing
the permeation of hydrated ions with a high water permeability
simultaneous
\cite{Cohen_Tanugi_2012,Suk_2014,Cohen_Tanugi_2014,Cohen_Tanugi_2015,O_Hern_2014,O_Hern_2015,Surwade_2015,Walker_2017,Ghosh_2018}. Nevertheless,
the fabrication of subnanometer porous 2D membranes over a large area
with atomic precision remains technically challenging
\cite{Suk_2014,Rollings_2016,O_Hern_2012,Wang_2017}. Moreover, a small
portion of excessively large pores can significantly diminish salt
rejection.  Very recently, a number of reports have explored the
effect of surface charges on the selectivity of ion transport through
graphene nanopores, as a consequence of the electrostatic interactions
between solvated ions and the charged functional groups attached on
graphene \cite{Rollings_2016,Surwade_2014}. However, it remains
controversial that if the surface charges can result in a degree of
salt rejection, as the membrane charge and potential cannot be
precisely quantified and are sensitive to the surroundings and
chemical treatment \cite{Li_2008}. The difficulty in turn hinders a
fundamental understanding of the interactions between the atomically
thin 2D membrane and the electrolyte solution. Here, we demonstrate
macroscopic salt rejection through a sheet of large-area porous
graphene using electrostatic gating, thereby deepening the
understanding of ion-graphene interaction and its implications on
nanoporous graphene-based desalination applications.

@@

\Fig{} \ref{fig:1}a depicts the experimental setup characterizing
ionic diffusion through a sheet of double layer porous graphene
(PG). Two reservoirs, namely the high-concentration reservoir (HCR)
containing electrolyte solution with molar concentration $c_0$, and
the low-concentration reservoir (LCR) containing de-ionized (DI)
water, are separated by a PG membrane supported by polycarbonate
track-etched (PCTE) film. Magnetic stirrers are used to minimize the
effects of external concentration polarization, and a conductivity
probe is placed in the LCR to monitor the ionic concentration as a
function of time. The membrane is attached to a piece of copper tape
connected to a voltage source applying an electrical bias
$V_{\mathrm{G}}$, with the LCR grounded (\Fig{} \ref{fig:1}a
inset). We fabricated the PCTE-supported PG using the protocol
developed by our group \cite{Choi_2018}, as schematically shown in
\Fig{} \ref{fig:1}b. The normally distributed, circular pores with
pore diameters of 20$\pm$10 nm and a pore density of
1.25$\pm$0.25$\times$10$^{10}$ cm$^{-2}$ (\Fig{} \ref{fig:1}c inset) over a
large area were perforated on graphene.  By optimizing the subsequent
transfer process, a high surface coverage ($>$ 98\%) of PG on PCTE was
obtained (\Fig{} \ref{fig:1}c), allowing reliable characterization of
the ion transport behavior (details see Supplementary Information
Section S1).  The system presented here allows us to systematically
investigate the diffusive flux of ions across a sheet of PG,
$\boldsymbol{J}$, driven by the gradient of electrochemical potential
in the electrolyte medium.

First, the control experiments were carried out by measuring the ionic
diffusion of seven salt species, including KCl, NaCl, LiCl,
K$_{2}$SO$_{4}$, MgSO4$_{4}$, CaCl$_{2}$ and K$_{3}$[Fe(CN)$_{6}$],
through the bare PCTE membrane at $c_{0}$ = 0.1 mM. As the diffusive
flux is comparably small, the conductivity varies linearly with time
(e.g. \Fig{} \ref{fig:2}a) within the measurement duration (1 h),
suggesting the concentration gradient across the membrane remains
constant. The diffusive fluxes of different salts were
obtained by converting the conductivity-time profiles with the
calibration curves (blue bars in \Fig{} \ref{fig:2}b and Supplementary
\Fig{} S2). For comparison purposes, based on the assumption that the
membrane consists cylindrical pores with uniform diameter and a low
tortuosity $\tau$ of $\sim{}$1.2 \cite{O_Hern_2012}, the molar
diffusive flux through the bare PCTE membrane,
$\boldsymbol{J}_{\mathrm{PCTE}}$, follows:
  \begin{equation}
    \label{eq:j-pcte}
    \boldsymbol{J}_{\mathrm{PCTE}} = \frac{\pi a^{2} N_{\mathrm{p}} D \Delta c}{\delta \tau}
  \end{equation}
where $a$ is the PCTE pore radius, $N_{\mathrm{p}}$ is numbers of pore
per area, $D$ is the salt diffusivity, and $\Delta c$ is the bulk
concentration difference between two reservoirs, i.e.,
$\Delta c \sim c_{0}$, $\delta$ is the thickness of the PCTE membrane,
and $\tau$ is the tortuosity of the PCTE membrane. By using the salt diffusivity values
provided by the PTCE vendor (Supplementary Table S1), the
calculated diffusive fluxes (red bars in \Fig{} \ref{fig:2}b) show
reasonable agreement with the experimental values.

Next, we carried out the same experiments using the PG-covered PCTE
membrane (PG-PCTE), without electrostatic gating. The measured
diffusive flux values, $\boldsymbol{J}_{\mathrm{PG0}}$ (green bars in
\Fig{} \ref{fig:2}b), remain at the same order of magnitude compared
with those of bare PCTE membrane, suggesting that the mass transfer
resistance of PG is not significant, as the permeability is high. We
observed a more pronounced decrease of diffusive flux in KCl,
K$_{2}$SO$_{4}$ and K$_{3}$[Fe(CN)$_{6}$] systems but not for the rest
of the salts. We hypothesized that the strong cation-$\pi$ interactions
between K$^{+}$ cations and graphene slow down the migration of cations
along the surface-mediated transport pathway \cite{Sun_2014}.

@@

We next discuss the effects of electrostatic gating on graphene. In
order to avoid any interference from the conductivity probe to the
membrane potential, the “three-interval” method, which has been used
in controlling the membrane potential in mesoporous carbon membranes
\cite{Surwade_2014}, is adopted here. \Fig{} \ref{fig:2}a
illustratively represents how the measured conductivity evolves with
time. Specifically, in the first interval, the gate voltage source is
turned off and the conductivity probe is on, allowing us to obtain the
conductivity-time profile with an average slope $s_{1}$ that
corresponds to the diffusive flux across the PG-PCTE membrane, . In
the second interval, a voltage $V_{\mathrm{G}}$ is applied to the
membrane after turning off the conductivity probe, followed by the
third interval $s_{3}$, in which the conductivity probe is turned on
again to give the corresponding slope. The slope corresponding to the
second interval $s_{2}$, as the conductivity probe is off, is
determined by linear interpolation of the conductivities from the
endpoint of interval 1 to the starting point of interval
3. Accordingly, we define the salt rejection factor, $\xi$, which is
effectively equivalent to the relative degree of diffusive flux
decrease as follows:
\begin{equation}
  \label{eq:rejection}
  \xi = \frac{\bar{s} - s_{2}}{\bar{s}} = \frac{\boldsymbol{J}_{\mathrm{PG0}}
    - \boldsymbol{J}_{\mathrm{PG}}(V_{\mathrm{G}})}{\boldsymbol{J}_{\mathrm{PG0}}}
\end{equation}
where $\bar{s}$ is the average slope of $s_{1}$ and $s_{3}$, namely
$(s_{1} + s_{3})/2$, and
$\boldsymbol{J}_{\mathrm{PG}}(V_{\mathrm{G}})$ is the diffusive flux
across the PTCE-supported PG double layer as a function of
$V_{\mathrm{G}}$. Note that a positive $\xi$ between 0 and 1
represents a decrease of the diffusive flux. The factor $\xi$ enables
a reliable and stable measurement of salt rejection, as the salt
species and initial conditions may induce a degree of measurement
uncertainty between different samples. A control experiment was
performed to ensure the copper tape was not in direct contact with the
ionic solution (Supplementary Section 3).

@@

In order to quantitatively understand the observed $V_{\mathrm{G}}$
dependence, we develop a theory to model the coupling of graphene’s
elementary transport properties and EDL, in order to quantify the
ionic transport through a graphene nanopore. Under the assumption that
the time scale for the bulk concentration change is significantly
longer than that of ionic diffusion, the pseudo-steady state
approximation holds. Accordingly, the steady-state Nernst-Planck
equation describing the ionic transport in an electrolyte solution is
given by \cite{MacGillivray_1968}:
\begin{equation}
  \label{eq:pnp}
  \nabla \cdot \boldsymbol{J}_{i} = -\nabla \cdot (\frac{D_{i}}{k_{\mathrm{B}}T} c_{i} N_{\mathrm{A}} \nabla \mu_{i}) = 0
\end{equation}
where $\boldsymbol{J}$ is the mass flux, subscript \textit{i}
corresponds to the ionic species (anion or cation), $D$ is the diffusivity,
$k_{\mathrm{B}}$ is the Boltzmann constant, $T$ is temperature, $c$ is
the molar concentration, $N_{\mathrm{A}}$ is the Avogadro constant,
and $\mu_{i}$ is the electrochemical potential. Under the assumption
of ideal solution, it follows \cite{Kilic_2007}:
\begin{equation}
  \label{eq:mu}
  \nabla \mu_{i} = k_{\mathrm{B}} T \nabla \ln x_{i} + z_{i} e \nabla \psi
\end{equation}
 $x$ is the molar fraction in solution, $z$ is the ionic valence,
$e$ is the unit charge, and $\psi$ is the electric potential. On the
other hand, the Poisson equation describing the electric potential
distribution is given by:
\begin{equation}
  \label{eq:poisson}
  \nabla \cdot (\varepsilon_{\mathrm{m}} \varepsilon_{0} \nabla \psi)
  =
  - N_{\mathrm{A}} e \sum_{i} c_{i} z_{i}
\end{equation}
where $\varepsilon_{\mathrm{m}}$ is the relative permittivity of
individual materials in the system, including water and the internal
Stern layer\cite{Tian_2017}, and
$\varepsilon_{0}$ is the vacuum permittivity. When applying
$V_{\mathrm{G}}$ to graphene adjacent to the electrolyte solution,
charges (electrons or holes) are induced in graphene; the
electroneutrality of the entire system suggests:
\begin{equation}
  \label{eq:electro-neutral}
  \sigma_{\mathrm{G}} S_{\mathrm{G}} + \sum_{i} \int_{\Omega} z_{i} c_{i} N_{\mathrm{A}} e \mathrm{d}^{3} \Omega= 0
\end{equation}
where $\sigma_{\mathrm{G}}$ is the charge density in graphene,
$S_{\mathrm{G}}$ is the total area of graphene, and $\Omega$
corresponds to the entire volumetric domain of electrolyte
solution. Note that the graphene surface potential,
$\psi_{\mathrm{G}}$, is not equivalent to $V_{\mathrm{G}}$ due to
change of graphene's work function upon charging, known as the
graphene quantum capacitance effect\cite{Xia_2009} (\Fig{}
\ref{fig:3}a) following:
\begin{equation}
  \label{eq:Vg}
  V_{\mathrm{G}} = \Delta \phi_{\mathrm{G}} + \psi_{\mathrm{G}}
\end{equation}

@@

$\Delta \phi_{\mathrm{G}} = \phi_{\mathrm{G}}(V_{\mathrm{g}}) -
\phi_{\mathrm{G}}(V_{\mathrm{g}}=0)$ is the change of graphene’s work
function. The properties of the PG samples fabricated is close to
intrinsic graphene, with charge density at $V_{\mathrm{G}}=0$ measured
to be XXX $e \cdot$cm$^{-2}$, corresponding to a work function
difference of XXX meV from graphene's Dirac point (details see
Supplementary Section SXX), much smaller than the
$\Delta \phi_{\mathrm{G}}$ induced by gating.  Therefore we adapt the
simplification that the charge neutrality point of graphene coincides
with the Dirac point at $V_{\mathrm{G}}$ = 0. The elementary
electronic properties of graphene give:
\begin{equation}
  \label{eq:delta-phiG}
  \Delta \phi_{\mathrm{G}} = \mathrm{sign}(\sigma_{\mathrm{G}}) \frac{\hbar v_{\mathrm{F}}}{e}
  \sqrt{\frac{\pi |\sigma_{\mathrm{G}}|}{e}}
\end{equation}
where $\hbar$ is the reduced Planck constant, and
$v_{\mathrm{F}}=1.1\times10^{6}$ m$\cdot$s$^{-1}$ is the Fermi
velocity of graphene (more details find Supplementary Section S5).
Note that the PG film fabricated experimentally is double-layer
turbostratic (randomly aligned) graphene, in which we assume the
$\Delta \phi_{\mathrm{G}} - \sigma_{\mathrm{G}}$ dependence follows
Eq. \eqref{eq:delta-phiG}. The theory proposed here was solved
self-consistently by discretizing
Eqs. \eqref{eq:pnp}-\eqref{eq:poisson} using the finite-element method
(FEM), in which the graphene surface potential is coupled with
Eqs. \eqref{eq:electro-neutral}-\eqref{eq:delta-phiG} that were solved
simultaneously to reach convergent numerical solutions of
$\sigma_{\mathrm{G}}$ and $\psi_{\mathrm{G}}$ for a given
$V_{\mathrm{G}}$. Clearly, the above model yields symmetric
characteristics for $\sigma_{\mathrm{G}}$ and $\psi_{\mathrm{G}}$ with
respect to $V_{\mathrm{G}}$ due to the symmetric band structure of
graphene (Eq. \eqref{eq:delta-phiG}). However, for graphene
synthesized by chemical vapor deposition (CVD) used in our
experiments, it is well-recognized that the electron traps are
inherently introduced during synthesis and patterning
\cite{Dean_2010}, effectively reducing the charge density and surface
potential on graphene $V_{\mathrm{G}}<0$ (Supplementary \Fig{} S3),
which is related to the fact that observed $\xi$ is much lower within
the $V_{\mathrm{G}}<0$ regime in \Fig{} \ref{fig:2}c. With the
nonideality in mind, hereafter, we compare the calculations and
experiments in the regime of $V_{\mathrm{G}}>0$. For example, \Fig{}
\ref{fig:3}c presents the calculated $\sigma_{\mathrm{G}}$ and
$\psi_{\mathrm{G}}$ as a function of $V_{\mathrm{G}}$ considering the
ionic diffusion through a single 20 nm-diameter nanopore drilled on a
sheet of semi-infinite double layer graphene that separates HCR
containing KCl solution at $c_{0}$ = 0.1 mM, and LCR at $c_{0}/10$ in
axisymmetric coordinates (details see Supplementary Section
S5). Indeed, the interplay between graphene quantum capacitance and
the EDL significantly reduces the graphene surface potential
$\psi_{\mathrm{G}}$ to $\sim$0.3 V at $V_{\mathrm{G}}$ = 1.25 V,
corresponding a surface charge density $\sigma_{\mathrm{G}}$ of

>>>>

\author{Roman M. Wyss} \affiliation{Soft Materials,
  Department of Materials, Eidgenössische Technische
  Hochschule (ETH) Zürich, Vladimir-Prelog-Weg 1-5, Zürich CH-8093,
  Switzerland.  }
\altaffiliation{R. M. S. and T. T. contributed equally to this work}
\author{Tian Tian} \affiliation{Institute for
  Chemical and Bioengineering Department of Chemistry and Applied
  Biosciences, Eidgenössische Technische Hochschule (ETH) Zürich,
  Vladimir-Prelog-Weg 1-5, Zürich CH-8093, Switzerland.}
\altaffiliation{R. M. S. and T. T. contributed equally to this work}

@@

%%%%%%%%%%%%%%%%%%%%
%Main
%%%%%%%%%%%%%%%%%%%%
\begin{document}
\clearpage{}
\begin{abstract}
  {\normalsize Atomically thin porous graphene membranes are emerging as one
    of the most promising candidates as next-generation membrane
    technology owing to their unique transport physics and high permeation. Among the current challenges of 2D membranes
    is the understanding of fundamental ion transport as well as its control towards applications in ion rectification, ionic transistors or desalination. 
    Here we report the study of tailoring the ion transport 
    through atomically-thin porous graphene with pore sizes of
    20$\pm$10 nanometers, and a pore density of 1.25$\pm$0.25$\times$10$^{10}$
    cm$^{-2}$. By electrostatically gating a sheet of porous graphene, we demonstrate
    a considerable reduction of salt flux by up to 55\% and
    85\% using a gate voltage, based on our experiments and
    theoretical predictions, respectively. We systematically
    investigate the effects of salt concentration and species,
    including the development of a theory to model the electrolyte diffusion
    through a nanopore in a sheet of gated graphene. The interplay
    between graphene quantum capacitance and electrical double layer
    is identified to be the main mechanism responsible for the
    observed salt flux reduction, when the pore size is comparable to the
    Debye length. Our findings reveal a new degree of
    freedom controlling the membrane potential and electrolyte
    permeation through porous two-dimensional materials.  }
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Main text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage{}
\linenumbers
\section{Introduction}
\label{sec:intro}
A central objective in the emerging field of nanofluidics is to
control transport phenomena at nanometer scale to enable and study novel
devices and systems. The reduction of characteristic
dimensions in nanofluidics gives rise to substantially different
transport properties in comparison to their bulk counterparts
\cite{Schoch_2008}. In nanopores, for example, the effect of
surface-mediated transport becomes dominant when the pore size is
comparable to the characteristic length of diffusion, enabling new
applications such as ionic diodes \cite{Karnik_2007}, field-effect
transistors \cite{Nam_2009}, desalination \cite{Heiranian_2015}, and
nanopore-based DNA sequencing \cite{Heerema_2016,Garaj_2013}. Porous,
atomically thin two-dimensional (2D) materials, such as graphene, are
emerging as ideal membrane materials due to their ultimate permeation
\cite{Suk_2010,Jiang_2009,Celebi_2014,Koenig_2012,Drahushuk_2012}. Early
research has suggested that using nanoporous graphene as a
desalination membrane enables excellent salt rejection by preventing
the passage of ions, maintaining a high water permeability
simultaneously.
\cite{Cohen_Tanugi_2012,Suk_2014,Cohen_Tanugi_2014,Cohen_Tanugi_2015,O_Hern_2014,O_Hern_2015,Surwade_2015,Walker_2017,Ghosh_2018}. Nevertheless,
the fabrication of subnanometer porous 2D membranes over a large area
with atomic precision remains technically challenging
\cite{Suk_2014,Rollings_2016,O_Hern_2012,Wang_2017}. Moreover, a small
portion of excessively large pores can significantly diminish salt
rejection.  Very recently, a number of reports have explored the
effect of surface charges on the selectivity of ion transport through
graphene nanopores, as a consequence of the electrostatic interactions
between solvated ions and the charged functional groups attached on
graphene, indicating that even large pores ($>$ 30nm) can be ion selective \cite{Rollings_2016,Surwade_2014}. However, it remains
controversial if surface charges effectively allow controlled salt -transport 
or -rejection, as the membrane charge and potential cannot be
precisely quantified and are sensitive to the surroundings and
chemical treatment \cite{Li_2008}. Therefore, there is a lack of 
fundamental understanding of the interactions between the atomically
thin 2D membrane, its surface charge and the electrolyte solution. Here, we study 
macroscopic salt rectification through a sheet of large-area porous
graphene using electrostatic gating, thereby deepening the
understanding of ion-graphene interaction and sheding light on applications of
nanoporous graphene for tailored ion transport.

@@

\Fig{} \ref{fig:1}a depicts the experimental setup characterizing
ionic diffusion through a sheet of double layer porous graphene
(PG). Two reservoirs, namely the high-concentration reservoir (HCR)
containing electrolyte solution with molar concentration $c_0$, and
the low-concentration reservoir (LCR) containing de-ionized (DI)
water, are separated by a PG membrane supported by polycarbonate
track-etched (PCTE) film. Magnetic stirrers are used to minimize the
effects of external concentration polarization, and a conductivity
probe is placed in the LCR to monitor the ionic concentration as a
function of time. The membrane is attached to a piece of copper tape
connected to a voltage source applying an electrical bias
$V_{\mathrm{G}}$, with the LCR grounded (\Fig{} \ref{fig:1}a
inset). We fabricated the PCTE-supported PG using the protocol
developed by our group \cite{Choi_2018}, as schematically shown in
\Fig{} \ref{fig:1}b. The normally distributed, circular pores with
pore diameters of 20$\pm$10 nm and a pore density of
1.25$\pm$0.25$\times$10$^{10}$ cm$^{-2}$ (\Fig{} \ref{fig:1}c inset) over a
large area were perforated on graphene.  By optimizing the subsequent
transfer process, a high surface coverage ($>$ 98\%) of PG on PCTE was
obtained (\Fig{} \ref{fig:1}c), allowing reliable characterization of
the ion transport behavior (details see Supplementary 
Section S1).  The system presented here allows us to systematically
investigate the diffusive flux of ions across a sheet of PG,
$\boldsymbol{J}$, driven by the gradient of electrochemical potential
in the electrolyte medium.

First, the control experiments were carried out by measuring the ionic
diffusion of seven salt species, including KCl, NaCl, LiCl,
K$_{2}$SO$_{4}$, MgSO$_{4}$, CaCl$_{2}$ and K$_{3}$[Fe(CN)$_{6}$],
through the bare PCTE membrane at $c_{0}$ = 0.1 mM. As the diffusive
flux is comparably small, the conductivity increases linearly with time in the LCR
(e.g. \Fig{} \ref{fig:2}a) within the measurement duration (1 h),
suggesting the concentration gradient across the membrane remains 
constant. The diffusive fluxes (blue bars in \Fig{} \ref{fig:2}b) of different salts were
obtained by converting the conductivity-time profiles using
calibration curves (Supplementary Section S2). The theoretical diffusive fluxes through PCTE are obtained based on the assumption that the
membrane consists of cylindrical pores with uniform diameter and a low
tortuosity $\tau$ of $\sim{}$1.2 \cite{O_Hern_2012}:

  \begin{equation}
    \label{eq:j-pcte}
    \boldsymbol{J}_{\mathrm{PCTE}} = \frac{\pi a^{2} N_{\mathrm{p}} D \Delta c}{\delta \tau}
  \end{equation}

where $a$ is the PCTE pore radius, $N_{\mathrm{p}}$ is numbers of pore
per area, $D$ is the salt diffusivity (average of cation and anion diffusivity), $\Delta c$ is the bulk
concentration difference between two reservoirs, i.e.,
$\Delta c \sim c_{0}$ and $\delta$ is the thickness of the PCTE membrane.
 By using the salt diffusivity values (Supplementary Table S1), the
calculated diffusive fluxes (red bars in \Fig{} \ref{fig:2}b) show
reasonable agreement with the experimental values.

Next, we carried out the same experiments using the PG-covered PCTE
membrane (PG-PCTE), without electrostatic gating. The measured
diffusive flux values, $\boldsymbol{J}_{\mathrm{PG0}}$ (green bars in
\Fig{} \ref{fig:2}b), remain at the same order of magnitude compared
with those of bare PCTE membrane, suggesting that the mass transfer
resistance of PG is not significant, as the permeability is high. 
The diffusive transport resistance through graphene pores scales with $\sim$ 1/D while
for a PCTE pore, the scaling follows $\sim$ L/D$^2$ \cite{O_Hern_2014},
 D being the respective diameter and L the channel length. 
Taking into account that there are 10 graphene pores
per PCTE pore on average, the transport resistance ratio of R$_{graphene}$/R$_{PCTE}$ is roughly 1/30,
 indicating that the PCTE dominates the overall ion transport, which is in line 
 with the observed behavior.

%Interestingly, we observed a pronounced decrease of diffusive flux in KCl,
%K$_{2}$SO$_{4}$ and K$_{3}$[Fe(CN)$_{6}$] systems but not for the rest
%of the salts. We hypothesized that the strong cation-$\pi$ interactions
%between K$^{+}$ cations and graphene slow down the migration of cations
%along the surface-mediated transport pathway \cite{Sun_2014}.

@@

We next discuss the effects of electrostatic gating on graphene. In
order to avoid any interference from the conductivity probe to the
membrane potential, the “three-interval” method, which has been used
in controlling the membrane potential in mesoporous carbon membranes
\cite{Surwade_2014}, is adopted here. \Fig{} \ref{fig:2}a
illustratively represents how the measured conductivity evolves with
time. Specifically, in the first interval, the gate voltage source is
turned off and the conductivity probe is on, allowing us to obtain the
conductivity-time profile with an average slope $s_{1}$ that
corresponds to the diffusive flux across the PG-PCTE membrane. In
the second interval, a voltage $V_{\mathrm{G}}$ is applied to the
membrane after turning off the conductivity probe, followed by the
third interval $s_{3}$, in which the conductivity probe is turned on
again to give the corresponding slope. The slope corresponding to the
second interval $s_{2}$, as the conductivity probe is off, is
determined by linear interpolation of the conductivities from the
endpoint of interval 1 to the starting point of interval
3. Accordingly, we define the salt rejection factor, $\xi$, which is
effectively equivalent to the relative degree of diffusive flux
decrease as follows:
\begin{equation}
  \label{eq:rejection}
  \xi = \frac{\bar{s} - s_{2}}{\bar{s}} = \frac{\boldsymbol{J}_{\mathrm{PG0}}
    - \boldsymbol{J}_{\mathrm{PG}}(V_{\mathrm{G}})}{\boldsymbol{J}_{\mathrm{PG0}}}
\end{equation}
where $\bar{s}$ is the average slope of $s_{1}$ and $s_{3}$, namely
$(s_{1} + s_{3})/2$, and
$\boldsymbol{J}_{\mathrm{PG}}(V_{\mathrm{G}})$ is the diffusive flux
across the PTCE-supported PG double layer as a function of
$V_{\mathrm{G}}$. Note that a positive $\xi$ between 0 and 1
represents a decrease of the diffusive flux. The factor $\xi$ enables
a reliable and stable measurement of salt rejection, as the salt
species and initial conditions may induce a degree of measurement
uncertainty between different samples. A control experiment was
performed to ensure the copper tape was not in direct contact with the
ionic solution (Supplementary Section 3).

@@

In order to quantitatively understand the observed $V_{\mathrm{G}}$
dependence, we develop a theory to model the coupling of graphene’s
elementary transport properties and EDL, in order to quantify the
ionic transport through a graphene nanopore. Under the assumption that
the time scale for the bulk concentration change is significantly
longer than that of ionic diffusion, the pseudo-steady state
approximation holds. Accordingly, the steady-state Nernst-Planck
equation describing the ionic transport in an electrolyte solution is
given by \cite{MacGillivray_1968}:
\begin{equation}
  \label{eq:pnp}
  \nabla \cdot \boldsymbol{J}_{i} = -\nabla \cdot (\frac{D_{i}}{k_{\mathrm{B}}T} c_{i} N_{\mathrm{A}} \nabla \mu_{i}) = 0
\end{equation}
where $\boldsymbol{J}$ is the mass flux, subscript \textit{i}
corresponds to the ionic species (anion or cation), $D$ is the diffusivity,
$k_{\mathrm{B}}$ is the Boltzmann constant, $T$ is temperature, $c$ is
the molar concentration, $N_{\mathrm{A}}$ is the Avogadro constant,
and $\mu_{i}$ is the electrochemical potential. Under the assumption
of ideal solution, it follows \cite{Kilic_2007}:
\begin{equation}
  \label{eq:mu}
  \nabla \mu_{i} = k_{\mathrm{B}} T \nabla \ln x_{i} + z_{i} e \nabla \psi
\end{equation}
 $x$ is the molar fraction in solution, $z$ is the ionic valence,
$e$ is the unit charge, and $\psi$ is the electric potential. On the
other hand, the Poisson equation describing the electric potential
distribution is given by:

\begin{equation}
  \label{eq:poisson}
  \nabla \cdot (\varepsilon_{\mathrm{m}} \varepsilon_{0} \nabla \psi)
  =
  - N_{\mathrm{A}} e \sum_{i} c_{i} z_{i}
\end{equation}

where $\varepsilon_{\mathrm{m}}$ is the relative permittivity of
individual materials in the system, including water and the internal
Stern layer\cite{Tian_2017}, and
$\varepsilon_{0}$ is the vacuum permittivity. When applying
$V_{\mathrm{G}}$ to graphene adjacent to the electrolyte solution,
charges (electrons or holes) are induced in graphene; the
electroneutrality of the entire system suggests:
\begin{equation}
  \label{eq:electro-neutral}
  \sigma_{\mathrm{G}} S_{\mathrm{G}} + \sum_{i} \int_{\Omega} z_{i} c_{i} N_{\mathrm{A}} e \mathrm{d}^{3} \Omega= 0
\end{equation}
where $\sigma_{\mathrm{G}}$ is the charge density in graphene,
$S_{\mathrm{G}}$ is the total area of graphene, and $\Omega$
corresponds to the entire volumetric domain of electrolyte
solution. Note that the graphene surface potential,
$\psi_{\mathrm{G}}$, is not equivalent to $V_{\mathrm{G}}$ due to
change of graphene's work function upon charging, known as the
graphene quantum capacitance effect\cite{Xia_2009} (\Fig{}
\ref{fig:3}a) following:

\begin{equation}
  \label{eq:Vg}
  V_{\mathrm{G}} = \Delta \phi_{\mathrm{G}} + \psi_{\mathrm{G}}
\end{equation}

@@

$\Delta \phi_{\mathrm{G}} = \phi_{\mathrm{G}}(V_{\mathrm{g}}) -
\phi_{\mathrm{G}}(V_{\mathrm{g}}=0)$ is the change of graphene’s work
function. The properties of the PG samples fabricated is close to
intrinsic graphene, with charge density at $V_{\mathrm{G}}=0$ measured
to be XXX $e \cdot$cm$^{-2}$, corresponding to a work function
difference of XXX meV from graphene's Dirac point (details see
Supplementary Section S5), much smaller than the
$\Delta \phi_{\mathrm{G}}$ induced by gating.  Therefore we adapt the
simplification that the charge neutrality point of graphene coincides
with the Dirac point at $V_{\mathrm{G}}$ = 0. The elementary
electronic properties of graphene give:

\begin{equation}
  \label{eq:delta-phiG}
  \Delta \phi_{\mathrm{G}} = \mathrm{sign}(\sigma_{\mathrm{G}}) \frac{\hbar v_{\mathrm{F}}}{e}
  \sqrt{\frac{\pi |\sigma_{\mathrm{G}}|}{e}}
\end{equation}
